---
title: "ML Project - Data Cleaning - Socio-economic Factors"
author: "Louise Gatty"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE, echo= FALSE}
rm(list = ls())
setwd("~/Desktop/Master/M1/S2/ML/project/code")

knitr::opts_chunk$set(echo = TRUE)

# Load libraries
library(tidyverse)
library(readxl)
library(janitor)
library(dplyr)
```

# Import datasets

```{r import}
# List of values to be considered as NA
valeurs_na <- c("N/A - division par 0", "N/A - résultat non disponible", "NA", "")

# Import datasets with specified NA values and skipping the first rows
demography <- read_excel("../data/raw/demographie.xlsx", 
                         skip = 3, 
                         na = valeurs_na)

economy <- read_excel("../data/raw/eco.xlsx", 
                      skip = 3, 
                      na = valeurs_na)

employment <- read_excel("../data/raw/emploi.xlsx", 
                         skip = 3, 
                         na = valeurs_na)

housing <- read_excel("../data/raw/logement.xlsx", 
                      skip = 3, 
                      na = valeurs_na)

mobilities <- read_excel("../data/raw/mobilites.xlsx", 
                         skip = 3, 
                         na = valeurs_na)

income <- read_excel("../data/raw/revenu.xlsx", 
                      skip = 3, 
                      na = valeurs_na)

transportation <- read_excel("../data/raw/transport.xlsx", 
                         skip = 3, 
                         na = valeurs_na)
```

The original dataset contains data for all the departments in France, but we are only interested in the departments of the Île-de-France region. Therefore, we will filter the dataset to keep only the rows where the "Code" column starts with 75, 91, 92, 93, 94, 95, 77, or 78.

```{r}
# Function to keep only the rows where "Code" starts with 75, 91, 92, 93, 94, 95, 77, 78
filter_departments <- function(df) {
  df_filtered <- df %>%
    filter(grepl("^(75|91|92|93|94|95|77|78)", Code))
  return(df_filtered)
}

# Filter the datasets to keep only the specified departments
demography_filtered <- filter_departments(demography)
economy_filtered <- filter_departments(economy)
employment_filtered <- filter_departments(employment)
housing_filtered <- filter_departments(housing)
mobilities_filtered <- filter_departments(mobilities)
income_filtered <- filter_departments(income)
transportation_filtered <- filter_departments(transportation)

# Remove the original datasets to free up memory
rm(demography, economy, employment, housing, mobilities, income, transportation)
```

The next step is to merge the filtered datasets into a single dataset. We will use the "Code" column as the key to merge the datasets.

```{r}
# Merge the filtered datasets using the "Code" column as the key (exclude the "Libellé" variable except in demography_filtered)
merged_data <- demography_filtered %>%
  left_join(economy_filtered %>% select(-`Libellé`), by = "Code") %>%
  left_join(employment_filtered %>% select(-`Libellé`), by = "Code") %>%
  left_join(housing_filtered %>% select(-`Libellé`), by = "Code") %>%
  left_join(mobilities_filtered %>% select(-`Libellé`), by = "Code") %>%
  left_join(income_filtered %>% select(-`Libellé`), by = "Code") %>%
  left_join(transportation_filtered %>% select(-`Libellé`), by = "Code")

# Export the merged dataset to a CSV file
write_csv2(merged_data, "../data/processed/socio_eco_indicators.csv")
```


### DRAFT FROM HERE ###

The next step is to check the percentage of NA values in each column of the filtered datasets.

```{r}
# Function to count the percentage of NA values in each column
pourcentage_na <- function(df) {
  na_count <- sapply(df, function(x) sum(is.na(x)))
  total_count <- nrow(df)
  na_percentage <- (na_count / total_count) * 100
  return(na_percentage)
}

# Calculate the percentage of NA values in the demography dataset
na_demography <- pourcentage_na(demography_filtered)
na_demography # 0 everywhere

# Calculate the percentage of NA values in the economy dataset
na_economy <- pourcentage_na(economy_filtered)
na_economy # 0 everywhere

# Calculate the percentage of NA values in the employment dataset
na_employment <- pourcentage_na(employment_filtered)
na_employment # 0 everywhere

# Calculate the percentage of NA values in the housing dataset
na_housing <- pourcentage_na(housing_filtered)
na_housing # ~37% for 3 variables

# Calculate the percentage of NA values in the mobilities dataset
na_mobilities <- pourcentage_na(mobilities_filtered)
na_mobilities # 0 everywhere

# Calculate the percentage of NA values in the income dataset
na_income <- pourcentage_na(income_filtered)
na_income # ~60% for 2 variables

# Calculate the percentage of NA values in the transportation dataset
na_transportation <- pourcentage_na(transportation_filtered)
na_transportation # 0 everywhere
```

Next step is to better understand the NAs in the housing and income datasets.

We drop the "Part des revenus d'activité" and "Taux de pauvreté (seuil à 60% du revenu médian)" variables from the income dataset as they have a high percentage of NA values.

```{r}
# Display rows with NA values in the housing dataset
housing_na <- housing_filtered %>%
  filter(if_any(
    .cols = contains("2015-2019"), 
    .fns = is.na
  )) %>%
  select(Code, contains("2015-2019"))
```
