---
title: "ML project - Data Cleaning - Energy Indicators"
author: "Louise Gatty"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE, echo= FALSE}
rm(list = ls())
setwd("~/Desktop/Master/M1/S2/ML/project/code")

knitr::opts_chunk$set(echo = TRUE)

# Load libraries
library(tidyverse)
library(readxl)
library(janitor)
library(dplyr)
```

# Import datasets

```{r import}
# List of values to be considered as NA
valeurs_na <- c("N/A - division par 0", "N/A - résultat non disponible", "NA", "")

# Import datasets with specified NA values and skipping the first rows
resources <- read_excel("../data/raw/ressources.xlsx", 
                          skip = 3, 
                          na = valeurs_na)

GES <- read_excel("../data/raw/GES.xlsx", 
                       skip = 3, 
                       na = valeurs_na)
```

The original dataset contains data for all the departments in France, but we are only interested in the departments of the Île-de-France region. Therefore, we will filter the dataset to keep only the rows where the "Code" column starts with 75, 91, 92, 93, 94, 95, 77, or 78.

```{r}
# Function to keep only the rows where "Code" starts with 75, 91, 92, 93, 94, 95, 77, 78
filter_departments <- function(df) {
  df_filtered <- df %>%
    filter(grepl("^(75|91|92|93|94|95|77|78)", Code))
  return(df_filtered)
}

# Filter the datasets to keep only the specified departments
resources_filtered <- filter_departments(resources)
GES_filtered <- filter_departments(GES)

# Remove the original datasets to free up memory
rm(resources, GES)
```

The next step is to merge the filtered datasets into a single dataset. We will use the "Code" column as the key to merge the datasets.

```{r}
# Merge the filtered datasets using the "Code" column as the key (exclude the "Libellé" once)
merged_data <- resources_filtered %>%
  left_join(GES_filtered %>% select(-Libellé), by = "Code")

# Remove "x000d" from column names
colnames(merged_data) <- gsub("x000d", "", colnames(merged_data))

# Export the merged dataset to a CSV file
write_csv2(merged_data, "../data/processed/energy_indicators.csv")
```

